---
title: "Untitled"
author: "Josh Erickson"
date: "June 16, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(tidyverse)
library(raster)
library(nhdplusTools)
library(GGally)
library(conflicted)
```


#Introduction  
This is an exploration into wetlands on the Ksanka Ranger District in Northwest Montana. We will use a basline layer (Fish and Wildlife Service Wetland Mapper) as the target response and remote sensing grids as predictors. Within the response we will only target "Freshwater Forested/Shrub Wetland", "Freshwater Pond" and "Freshwater Emergent Wetland." We will then train the data using a multinomial Random Forest algorithm. Below is the workflow to get from pre-processing to model building.

Bring in polygons from the sample space and wetland mapper.

```{r}

proj_crs <- "+proj=aea +lat_1=46 +lat_2=48 +lat_0=44 +lon_0=-109.5 +x_0=600000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"

#point to some folders
hydro_folder <- "D:/documents/Documents/Josh/Documents/GIS/District_H20_exports/Hydro"
soils_folder <- "D:/documents/Documents/Josh/Documents/GIS/District_H20_exports/Soils"
lands_folder <- "D:/documents/Documents/Josh/Documents/GIS/District_H20_exports/Lands_District_stuff"
DEM_folder <- "D:/documents/Documents/Josh/Documents/GIS/District_H20_exports/DEM_Rasters"
HWpred_folder <- "D:/Rcodes/Water_Prediction/Hird_Water_Prediction/waterD/waterPred/Final_workflow"
prd_folder <- "D:/Rcodes/Water_Prediction/Final_paper_water_prediction/hwstreamprediction"

#sample space is the ranger district

district_boundary <- read_sf(paste0(lands_folder, "/district_bdy.shp")) %>%  st_transform(., proj_crs) %>% filter(., DISTRICT_N %in%  c("Fortine", "Rexford")) %>% select(OBJECTID)

district_boundary <- st_union(district_boundary)

district_boundary <- st_as_sf(district_boundary)


#these are the wetlands within the ranger district

wetland_HUC_17010106 <- read_sf(dsn = "HU8_17010106_watershed/HU8_17010106_Watershed/HU8_17010106_wetlands.shp")  %>%  st_transform(., proj_crs)%>% st_intersection(., district_boundary)

wetland_HUC_17010210 <- read_sf(dsn = "HU8_17010210_watershed/HU8_17010210_Watershed/HU8_17010210_wetlands.shp")  %>%  st_transform(., proj_crs)

wetland_HUC_17010101 <- read_sf(dsn = "HU8_17010101_watershed/HU8_17010101_wetlands.shp")  %>%  st_transform(., proj_crs)


#here we combined the wetlands into one df

wetland_combine <- rbind(wetland_HUC_17010101, wetland_HUC_17010106, wetland_HUC_17010210) %>% filter(WETLAND_TY %in% c("Freshwater Forested/Shrub Wetland", "Freshwater Pond", "Freshwater Emergent Wetland"))


#now filter out the 'test' watershed 'Sunday Creek' and intersect with sample space
nhdPlushr <- get_nhdplushr(prd_folder, "nhdplus.gpkg",
layers = "WBDHU12", overwrite = TRUE, proj = "+proj=aea +lat_1=46 +lat_2=48 +lat_0=44 +lon_0=-109.5 +x_0=600000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0")

dist_wo_Sunday <- st_as_sf(nhdPlushr$WBDHU12)  %>% filter(Name != "Sunday Creek") %>% st_intersection(., district_boundary) %>% st_union(.)

wetlands_eureka <- st_intersection(wetland_combine, dist_wo_Sunday)

#now get Sunday by itself
dist_with_Sunday <- st_as_sf(nhdPlushr$WBDHU12)  %>% filter(Name == "Sunday Creek")

#then grab the wetland data

wetlands_sunday <- st_intersection(wetland_combine, dist_with_Sunday)
ggplot() + geom_sf(data=wetlands_sunday, aes(fill = WETLAND_TY))

```

```{r}

#good we don't have any points in the Sunday Creek watershed

# now plot the wetlands

plot(wetlands_eureka[1])

```

Next step would be to add the points at random onto the sample space. This will allow us to randomly extract predictor values randomly.

```{r}
wetland_points <- st_sample(wetlands_eureka, 5000, "random") %>% st_as_sf(.)

land_points <- st_sample(dist_wo_Sunday, 5000, "random") %>% st_as_sf(.)

sample_points <- rbind(land_points, wetland_points) %>% mutate(id = row_number())

#now intersect with wetland data

wetland_pts <- st_intersection(wetlands_eureka, sample_points)

#then get the other points by id

land_pts <- anti_join(sample_points %>% as.data.frame(), wetland_pts %>% as.data.frame(),  by = "id") %>% mutate(WETLAND_TY = "No Wetland") %>% st_sf(sf_column_name = 'x') %>% rename(geometry = "x")

#now join both wetland_pts and land_pts

wetland_pts <- wetland_pts[,c(2,6)]

sample_pts <- rbind(wetland_pts, land_pts)

sample_pts <- sample_pts %>% rename(response = "WETLAND_TY")

ggplot() + geom_sf(data = sample_pts, aes(color = response))



```

```{r}
sample_pts %>% count(response)
sample_pts %>% count(response) %>% summarize(prop = paste((n/sum(n))*100, "%"), name = paste(response))

#might be an issue with CV and and the 5.53% for freshwater pond


```


Now we can create a raster stack to be able to extract predictor values from the points.

```{r}

#point to some folders
dem_folder <- "D:/Rcodes/Exploring/Cabinet/Soils/exploring_units"
HWpred_folder <- "D:/Rcodes/Water_Prediction/Hird_Water_Prediction/waterD/waterPred/Final_workflow"

```


Read in the Rasters

```{r}

#use TWI as the base raster to resample to, no reason could pick any.
twi <- raster(paste0(HWpred_folder, "/twi.tif"))

dem10 <- raster(paste0(dem_folder, "/dem10m.tif"))
slope <- crop(dem10, district_boundary) 
slope <- terrain(slope, opt = "slope", unit = "degrees")
slope <- resample(slope, twi, method = "ngb")
writeRaster(slope, "slope_wetland.tif", overwrite = TRUE)

TPI <- crop(dem10, district_boundary) 
TPI <- terrain(TPI, opt = "TPI")
TPI <- resample(TPI, twi, method = "ngb")
writeRaster(TPI, "TPI_wetland.tif", overwrite = TRUE)

TRI <- crop(dem10, district_boundary) 
TRI <- terrain(TRI, opt = "TRI")
TRI <- resample(TRI, twi, method = "ngb")
writeRaster(TRI, "TRI_wetland.tif", overwrite = TRUE)


aspect <- crop(dem10, district_boundary) 
aspect <- terrain(aspect, opt = "aspect")
aspect <- resample(aspect, twi, method = "ngb")
writeRaster(aspect, "aspect_wetland.tif", overwrite = TRUE)

ndvi <- raster(paste0(HWpred_folder, "/ndviAS.tif"))
ndvi <- resample(ndvi, twi, method = "ngb")
writeRaster(ndvi, "ndvi_wetland.tif", overwrite = TRUE)

ndwi <- raster(paste0(HWpred_folder, "/ndwiAS.tif"))
ndwi <- resample(ndwi, twi, method = "ngb")
writeRaster(ndwi, "ndwi_wetland.tif", overwrite = TRUE)

npol <- raster(paste0(HWpred_folder, "/npol.tif"))
npol <- resample(npol, twi, method = "ngb")
writeRaster(npol, "npol_wetland.tif", overwrite = TRUE)

vv <- raster(paste0(HWpred_folder, "/vvtesting.tif"))
vv <- resample(vv, twi, method = "ngb")
writeRaster(vv, "vv_wetland.tif", overwrite = TRUE)

vvsd <- raster(paste0(HWpred_folder, "/vvsd.tif"))
vvsd <- resample(vvsd, twi, method = "ngb")
writeRaster(vvsd, "vvsd_wetland.tif", overwrite = TRUE)

B2 <- raster(paste0(HWpred_folder, "/B2__81_930_15_19.tif"))
dist_longlat <- st_transform(district_boundary, crs = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
B2 <- crop(B2, dist_longlat) 
B2 <- projectRaster(B2, twi, res = 10, crs = proj_crs, method = "ngb")
writeRaster(B2, "b2_testing.tif", overwrite = TRUE)

B3 <- raster(paste0(HWpred_folder, "/B3_81_930_15_19.tif"))
B3 <- crop(B3, dist_longlat) 
B3 <- projectRaster(B3, twi, res = 10, crs = proj_crs, method = "ngb")
writeRaster(B3, "b3_testing.tif", overwrite = TRUE)

B4 <- raster(paste0(HWpred_folder, "/B4_81_930_15_19.tif"))
B4 <- crop(B4, dist_longlat) 
B4 <- projectRaster(B4, twi, res = 10, crs = proj_crs, method = "ngb")
writeRaster(B4, "b4_testing.tif", overwrite = TRUE)

NIR <- raster(paste0(HWpred_folder, "/B8_81_930_15_19.tif"))
NIR <- crop(NIR, dist_longlat) 
NIR <- projectRaster(NIR, twi, res = 10, crs = proj_crs, method = "ngb")
writeRaster(NIR, "NIR_testing.tif", overwrite = TRUE)
```

Take the rasters and make a stack.

```{r}

wetland_stack <- stack(NIR, B4, B3, B2, vvsd, vv, npol, ndwi, ndvi, TPI, TRI, twi, slope, aspect)
```



Now we need to extract the values from the raster stack


```{r}

data <- extract(wetland_stack,sample_pts)

data <- data.frame(cbind(data, sample_pts$response))

#now create an SF data frame for visualizing

dataSF <- data.frame(data, sample_pts)

dataSF <- dataSF %>% select(-c("V15", "id"))

```

Before we move on let's rename some of the columns and clean up the dataframe.

```{r}
data <- data %>% rename(NIR = "B8_81_930_15_19", Red = "B4_81_930_15_19", Green = "B3_81_930_15_19", Blue = "B2__81_930_15_19", response = "V15") %>% select(response, everything()) %>% mutate(across(-response, as.numeric))%>% na.omit()

dataSF <- dataSF %>% rename(NIR = "B8_81_930_15_19", Red = "B4_81_930_15_19", Green = "B3_81_930_15_19", Blue = "B2__81_930_15_19") %>% select(response, everything()) %>% mutate(across(-c("response","geometry"), as.numeric)) %>% mutate(response = as.factor(response)) %>% na.omit()

dataSF <- dataSF %>% na.omit()
```

```{r}
data_cor <- data %>% select(-response) %>% cor(.)
data_cor
corrplot::corrplot(data_cor, "ellipse")

# or 

ggcorr(data[2:15]) + coord_flip()
```

Now look at the distributions

```{r}
data %>% pivot_longer(cols = -response, names_to = "rasters") %>% ggplot() + geom_density(aes(value, ..density..), adjust = 5) + scale_y_continuous(labels = scales::comma) + facet_wrap(~rasters, scales = "free_y")
```

Blue, Green, Red and NIR look like they have some long tails so we can check them out as well.

```{r, message=FALSE}
data %>% select(response, Blue, Green, Red, NIR) %>% pivot_longer(cols = -response, names_to = "rasters") %>% ggplot() + geom_boxplot(aes(value)) + facet_wrap(~rasters, scales = "free_x") + coord_flip()
```

Actually doesn't look too bad. We can also take a look at a `pairs` plot and see if the predictors have some relationships with the response in a multvariate fashion by coloring the response. 

```{r, message=FALSE}

#all the *terrain* type variables
data %>% sample_n(1000) %>%  ggpairs(aes(color = response), columns = c(1,11:15),
upper = list(continuous = wrap("density", alpha = 0.5, size = 1.5)),
lower = list(continuous = wrap("points", alpha = 0.3))
)

data %>% sample_n(1000) %>%  ggpairs(aes(color = response), columns = 1:10,
upper = list(continuous = wrap("density", alpha = 0.5, size = 1.5)),
lower = list(continuous = wrap("points", alpha = 0.3))
)

```

